US CES Analysis for Victor Court

### 2018-07-11 Articles/Court_Analysis_2018/FittingForCourt2018.Rmd
Edits to Victor Court analysis.

---
title: "Analysis for Victor Court"
author: "Matthew Kuperus Heun"
date: "10 July 2018"
output: pdf_document
---

@@

Victor has asked me to use my CES fitting functions on his US data
as a point of comparison.

@@

```{r}
USData_raw <- read.csv(file = "CES_USA.csv", sep = ";")
head(USData_raw, n = 3)
```

@@

```{r}
USData <- USData_raw %>% 
  select(-K.L...US.2010.per.hour., 
         -Capital...US.2010., 
         -Labor..millions.of.hours., 
         -Primary.exergy..TJ.,
         -GDP....US.2010.) %>% 
  rename(
    k = K..1960...1.,
    l = L..1960...1.,
    e = E..1960...1.,
    y = Y..1960...1.
  ) %>% 
  mutate(
    Country = "US"
  ) %>% 
  select(Country, Year, y, k, l, e)
head(USData, n = 3)
```

Now review the data.

@@

We can query which boundary model we have: 
```{r}
kleModel$bname
```
So this is boundary model 7 of Table 2 in the ``Cautionary Tales'' paper.

@@

With $\theta = \mathrm{e}^{\mathrm{logscale}} = \mathrm{e}^
{`r naturalCoef(kleModel)[["logscale"]]`} =
`r naturalCoef(kleModel)[["scale"]]`$,
these results mean that boundary model 7

@@

$$y = `r exp(coef(kleModel)[["logscale"]])` 
      \mathrm{e}^{`r coef(kleModel)[["lambda"]]` (t-1960)} 
      \left[ `r coef(kleModel)[["delta_1"]]` \, k 
          + `r 1 - coef(kleModel)[["delta_1"]]` \, l \right] \; .$$

@@

```{r}
USData %>% 
  mutate(
    `y predicted` = yhat(kleModel) %>% as.numeric() # avoids an attributes warning
  ) %>% 
  select(Year, y, `y predicted`) %>% 
  gather(key = var, value = value, y, `y predicted`) %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = var)) +
  geom_line() + 
  geom_hline(yintercept = 1, colour = "gray50", size = 0.1) +
  labs(x = NULL, y = NULL, linetype = NULL) +
  xy_theme()
```

@@

1. The fit is rather good.
2. It is likely that the fitted model excludes energy ($e$), 
   because the energy dip in 1980 puts the energy trajectory at a very different
   multiple of the $y$ curve. 
   In contrast, the $l$ curve stays at about the same multiple of $y$
   throughout the entire time series. 
   Thus, the $l$ curve . 
   Instead, by picking up the smaller dip in the $l$ data series, the fit is better.
3. I find it interesting, but don't quite know what to make of it, that
   the coefficients on $k$ and $l$ are approximately the neoclassical numbers 
   (0.7 and 0.3, respectively).

>>>>

---
title: "US CES Analysis for Victor Court"
author: "Matthew Kuperus Heun"
date: "10 July 2018"
output: pdf_document
---

@@

Victor Court has asked me to fit US energy and economic data
using the CES aggregate production function. 
The energy and economic data have been supplied by Victor.

@@

```{r}
USData_raw <- read.csv(file = "CES_USA.csv", sep = ";")
head(USData_raw, n = 3) # Show the first 3 rows
```

@@

```{r}
USData <- USData_raw %>% 
  select(-K.L...US.2010.per.hour., 
         -Capital...US.2010., 
         -Labor..millions.of.hours., 
         -Primary.exergy..TJ.,
         -GDP....US.2010.) %>% 
  rename(
    k = K..1960...1.,
    l = L..1960...1.,
    e = E..1960...1.,
    y = Y..1960...1.
  ) %>% 
  mutate(
    Country = "US"
  ) %>% 
  select(Country, Year, y, k, l, e) # Re-order the columns
head(USData, n = 3)
```

Now review the data in graphical form.

@@

We can query which boundary model we have obtained: 
```{r}
kleModel$bname
```
So this is boundary model 7 of Table 2 in the ``Cautionary Tales'' paper.

@@

With $\theta = \mathrm{e}^{\mathrm{logscale}} = \mathrm{e}^
{`r nckle[["logscale"]]`} = `r nckle[["scale"]]`$,
these results mean that boundary model 7

@@

$$y = `r nckle[["scale"]]` 
      \mathrm{e}^{`r coef(kleModel)[["lambda"]]` (t-1960)} 
      \left[ `r coef(kleModel)[["delta_1"]]` \, k 
          + `r 1 - coef(kleModel)[["delta_1"]]` \, l \right] \; .$$

At the moment, I'm a bit unclear about the $\theta$ coefficient.
Inserting values for 1960 ($k = 1$, $l = 1$)
gives $y = \theta \mathrm{e}^0 [1] = \theta = `r nckle[["scale"]]` \ne 1$,
but it should be about 1.

I probably have something wrong in my thinking about this.  
Any ideas?

@@

```{r}
USData %>% 
  mutate(
    `y fitted` = yhat(kleModel) %>% as.numeric() # avoids an attributes warning
  ) %>% 
  select(Year, y, `y fitted`) %>% 
  gather(key = var, value = value, y, `y fitted`) %>% 
  ggplot(mapping = aes(x = Year, y = value, linetype = var)) +
  geom_line() + 
  geom_hline(yintercept = 1, colour = "gray50", size = 0.1) +
  labs(x = NULL, y = NULL, linetype = NULL) +
  xy_theme()
```

@@

1. Visually, the fit is quite good.
2. It is likely that the fitted model excludes energy ($e$), 
   because the energy dip in 1980 puts the energy trajectory at a very different
   multiple of the $y$ curve before and after 1980. 
   Mixing $e$ into the fit makes the fitted model worse!
   In contrast, the $l$ curve stays at about the same multiple of $y$
   throughout the entire time series, yet retains the dip in 1980. 
   The $l$ curve provides a better approximation to the shape of $y$.
   So, the fitting algorithm chooses to use $l$ instead of $e$ to fit $y$.
3. I find it interesting, but don't quite know what to make of it, that
   the coefficients on $k$ and $l$ are approximately the neoclassical numbers 
   (0.7 and 0.3, respectively).
   Maybe it's a result of the fact that the $k$ and $l$ lines are
   approximately a ratio of 2 apart from each other.
